---
title: "Reproducable Research Week 4 Project 2 Report"
author: "Adam Weissman"
date: "8 September 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE)
```

## Overview
The basic goal of this assignment is to explore the NOAA Storm Database and answer some basic questions about severe weather events.

Documentation on the National Oceanic and Atmospheric Administration Storm dataset:
[National Weather Service Storm Data Documentation](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf)
[National Climatic Data Center Storm Events FAQ](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2FNCDC%20Storm%20Events-FAQ%20Page.pdf)

Setup dependency libraries, download the dataset, and load it into R.

```{r init}
library(knitr)
library(ggplot2)
library(dplyr)
library(readr)
library(lubridate)

dlzip="StormData.csv.bz2"
dlurl="https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
datafilename="StormData.csv.bz2"

if(!file.exists(dlzip)){
    download.file(dlurl,dlzip)
}
```
```{r readdata}
storm<-read_csv(dlzip)

```

## Clean data
Date field is fine, but want to convert to lubridate type. Time zones aren't in Olsen Names, but don't really need time, so I won't clean it.

```{r clean}
storm$BGN_DATE<-mdy_hms(storm$BGN_DATE)
storm$END_DATE<-mdy_hms(storm$END_DATE)
```

## Question 1 - Event Type Damage
Across the United States, which types of events (as indicated in the \color{red}{\verb|EVTYPE|}EVTYPE variable) are most harmful with respect to population health?

To find insight on this question with the NOAA storm data, a plot must be generated showing effects of each storm type on population health.

```{r q1}
popHealth<-storm %>% group_by(EVTYPE) %>% summarise(fatalWInj=sum(FATALITIES,INJURIES))
popHealth$EVTYPE<-with(popHealth,factor(popHealth$EVTYPE,levels=popHealth[order(-fatalWInj),]$EVTYPE))
popHealth<-arrange(popHealth,desc(fatalWInj))
plotData<-popHealth[1:10,]
plotData<-rbind.data.frame(plotData,cbind(EVTYPE="All other", fatalWInj=sum(popHealth[11:nrow(popHealth),2])))
g<-ggplot(plotData,aes(x=EVTYPE,y=as.numeric(fatalWInj)))
g+geom_col()+theme(axis.text.x=element_text(angle=90,hjust=1))+labs(title="Top 10 Most Dangerous Weather Event Types", x="Weather Event Types",y="Total Count of Fatalities and Injuries")
```

As shown in the plot, tornados are by far the 
